### 2022-10-01 aims-template/Template-AIMS Mathematics.tex
AIMS Mathematics template

\documentclass{aims}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{txfonts}
\def\typeofarticle{Type of article}
\def\currentvolume{7}
\def\currentissue{x}
\def\currentyear{2022}
\def\currentmonth{Received date 1 July 2016, Accepted date 6 September}
\def\ppages{xxx--xxx}
\def\DOI{10.3934/math.2022xxx}
\def\Received{July 2021}
\def\Revised{August 2021}
\def\Accepted{September 2021}
\def\Published{December 2021}

\numberwithin{equation}{section}
\DeclareMathOperator*{\essinf}{ess\,inf}
\makeatletter
\renewcommand{\@biblabel}[1]{#1\hfill \hspace{-0.2cm}}
\makeatother

\newcommand{\ep}{\varepsilon}
\newcommand{\eps}[1]{{#1}_{\varepsilon}}

\begin{document}

\title{The full title of your paper}

\author{%
  First name Last name\affil{1}
  First name Last name\affil{2}
  and
  First name Last name\affil{1,}\corrauth
}

% \shortauthors is used in copyright information in the end of the paper
\shortauthors{the Author(s)}

\address{%
  \addr{\affilnum{1}}{Department of Mathematics, Missouri State University, 901 S. National, Springfield, MO 65897, USA}
  \addr{\affilnum{2}}{Affiliation}}

% corresponding author
\corraddr{Email address of corresponding author; Tel: +1-111-111-1111;\\
 Fax: +1-111-111-1111.
}

\begin{abstract}
An abstract is a brief of the paper; the abstract should not 
contain references, the text of the abstract section should be
in 12 point normal Times New Roman.
\textbf{(200 to 300 words)}
\end{abstract}

\keywords{
\textbf{(5 to 10 keywords)}
\newline
\textbf{Mathematics Subject Classification:} xxx, xxx}

\maketitle

\section{Introduction}

Please use the AIMS template to prepare your manuscript, 
before you submit to our journal. 
Please read carefully the instructions for authors at
http://www.aimspress.com \cite{authour1}. 
These are important instructions and explanations. 
Thank you for your cooperation.

\section{Materials and methods}
\subsection{Subheading}

\subsubsection{Sub-subheading}
The heading levels should not be more than 4 levels. 
The font of heading and subheadings should be 12 point 
normal Times New Roman. The first letter of headings 
and subheadings should be capitalized.

\section{Results}
The body text is in 12 point normal Times New Roman, 
the line space is at least 15 point.

\begin{table}[H]
\begin{center}
\caption{Caption of the table.}
\begin{tabular}{ccc} \hline
 & & \\\hline
 & & \\
 & & \\
 & & \\\hline
 &(Table body should be created by MS word table function; three-line table is preferred.)
\end{tabular}
\end{center}
\end{table}

\begin{figure}[H]
\begin{center}
\includegraphics[scale=0.8]{figure.pdf}
\caption{Legend of the figure.}
\label{Fig1}
\end{center}
\end{figure}

\begin{equation}
  \text{[add an equation here; use MS Word or MathType equation function]}
\end{equation}

\section{Discussion}

\section{Conclusions}

\section*{Acknowledgments (All sources of funding of the study must be disclosed)}
We would like to thank you for following the instructions above 
very closely in advance. It will definitely save us lot of 
time and expedite the process of your paper's publication.

\section*{Conflict of interest}

\begin{thebibliography}{999}

\bibitem{authour1}[10.1090/S0894-0347-1992-1124979-1]
      \textbf{Journal article style:} Y. Benoist, P. Foulon, F. Labourie,  {Anosov flows with stable and unstable differentiable distributions},
     \newblock  {\it J. Amer. Math. Soc.,}
     \newblock  \textbf{Volume} (year), StaringPage-Ending Page. \doilink{http://dx.doi.org/10.1090/S0894-0347-1992-1124979-1}

\bibitem{authour2}[10.1016/B978-0-12-775850-3.50017-0]
     \textbf{Book style:}
    \newblock J. Serrin, {\em Gradient estimates for solutions of nonlinear elliptic and parabolic equations}, 
    \newblock  2 Eds., New York: Academic Press, 1971. \doilink{http://dx.doi.org/10.1016/B978-0-12-775850-3.50017-0}

\bibitem{authour3}
     \textbf{Online content:}
    \newblock  {\it SARS Expert Committee, SARS in Hong Kong: From Experience to Action}, Hong Kong SARS Expert Committee, 2003. Available from: \\ \url{http://www.sars-expertcom.gov.hk/english/reports/reports.html.}

\end{thebibliography}

For more questions regarding reference style, please refer to the \href{http://www.ncbi.nlm.nih.gov/books/NBK7256/}{Citing Medicine}.

\section*{Supplementary (if necessary)}

\end{document}