### 2022-10-04 aims-template/Template-AIMS Mathematics.tex
switch template

\documentclass{aims}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{txfonts,verbatim}

\numberwithin{equation}{section}
\DeclareMathOperator*{\essinf}{ess\,inf}
\makeatletter
\renewcommand{\@biblabel}[1]{#1\hfill \hspace{-0.2cm}}
\makeatother

\newcommand{\ep}{\varepsilon}
\newcommand{\eps}[1]{{#1}_{\varepsilon}}

\begin{document}

\title{The full title of your paper}

\author{%
  First name Last name\affil{1}
  First name Last name\affil{2}
  and
  First name Last name\affil{1,}\corrauth
}

% \shortauthors is used in copyright information in the end of the paper
\shortauthors{the Author(s)}

\address{%
  \addr{\affilnum{1}}{Department of Mathematics, Missouri State University, 901 S. National, Springfield, MO 65897, USA}
  \addr{\affilnum{2}}{Affiliation}}

% corresponding author
\corraddr{Email address of corresponding author; Tel: +1-111-111-1111;\\
 Fax: +1-111-111-1111.
}

\begin{abstract}
An abstract is a brief of the paper; the abstract should not 
contain references, the text of the abstract section should be
in 12 point normal Times New Roman.
\textbf{(200 to 300 words)}
\end{abstract}

\keywords{
\textbf{(5 to 10 keywords)}
\newline
\textbf{Mathematics Subject Classification:} xxx, xxx}

\maketitle
\end{document}